---
title: "appendix_stuff"
output: html_document
---

```{r}
#load packages
library(pacman)
pacman::p_load("tidyverse", "lubridate", "zoo", "strucchange")

```

```{r}
getwd()
```


```{r}
# load dataset
inf_data = read_csv("../../data/inference/predicted_inference_data_with_gov.csv")

#Make date column into date object
inf_data$date = as.Date(inf_data$date)
# 
#Get the month
inf_data$month <- as.yearmon(inf_data$date)
# Extract only the year
inf_data$year <- as.numeric(format(inf_data$month, "%Y"))

inf_data$party = as.factor(inf_data$party)
```
```{r}
inf_data
```


```{r}
library(dplyr)
library(zoo)
library(lubridate)

ts_data <- inf_data %>%
  mutate(month = as.yearmon(date)) %>%
  group_by(month) %>%
  summarise(n_sentences = n()) %>%
  arrange(month)


library(ggplot2)

countplot_1 <- ggplot(ts_data, aes(x = month, y = n_sentences)) +
  geom_line(color = "grey") +
  geom_smooth(se = FALSE, color = "#2C7BB6") +
  geom_vline(xintercept = as.yearmon("2019-01"),
             linetype = "dashed", color = "#C0392B", linewidth = 1) +
  #labs(title = "Sentence Counts pr Month",
       #subtitle = "No abrupt shift at dataset merge (2019)")+
       ylab("Sentences per Month")+
       xlab("Month")+
  theme_minimal()+
  theme(
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = "grey92", linewidth = 0.4),
    axis.line.x        = element_line(color = "grey80", linewidth = 0.4),
    axis.ticks.x       = element_line(color = "grey80", linewidth = 0.4),
    axis.ticks.length  = unit(3, "pt"),
    plot.title         = element_blank(),
    axis.title         = element_text(size = 15, color = "grey40"),
    axis.text          = element_text(size = 11,  color = "grey50")
  )
countplot_1
ggsave("countplot_1.png",countplot_1, width = 5, height = 4, dpi = 600)
```


#rerun with normalized for activity


```{r}

ts_data <- inf_data %>%
  mutate(month = as.yearmon(date)) %>%
  group_by(month) %>%
  summarise(
    n_sentences = n(),
    n_days = n_distinct(date)
  ) %>%
  mutate(sentences_per_day = n_sentences / n_days)

# countplot_2 <- ggplot(ts_data, aes(x = month, y = sentences_per_day)) +
#   geom_line(color = "grey") +
#   geom_smooth(se = FALSE, color = "#2C7BB6") +
#   geom_vline(xintercept = as.yearmon("2019-01"),
#              linetype = "dashed", "#C0392B", linewidth = 1) +
#          ylab("nr Monthly Sentences pr active day")+
#        xlab("Month")+
#   theme_minimal()+
#   theme(
#     panel.grid.minor   = element_blank(),
#     panel.grid.major.x = element_blank(),
#     panel.grid.major.y = element_line(color = "grey92", linewidth = 0.4),
#     axis.line.x        = element_line(color = "grey80", linewidth = 0.4),
#     axis.ticks.x       = element_line(color = "grey80", linewidth = 0.4),
#     axis.ticks.length  = unit(3, "pt"),
#     plot.title         = element_blank(),
#     axis.title         = element_text(size = 15, color = "grey40"),
#     axis.text          = element_text(size = 13,  color = "grey50")
#   )
# 
# countplot_2
# ggsave("countplot_2.png",countplot_2, width = 5, height = 4)

countplot_2 <- ggplot(ts_data, aes(x = month, y = sentences_per_day)) +
  geom_line(color = "grey") +
  geom_smooth(se = FALSE, color = "#2C7BB6") +
  geom_vline(xintercept = as.yearmon("2019-01"),
             linetype = "dashed", color = "#C0392B", linewidth = 1) +
  #labs(title = "Sentence Counts pr Month",
       #subtitle = "No abrupt shift at dataset merge (2019)")+
       ylab("Monthly Sentences per Active Day")+
       xlab("Month")+
  theme_minimal()+
  theme(
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = "grey92", linewidth = 0.4),
    axis.line.x        = element_line(color = "grey80", linewidth = 0.4),
    axis.ticks.x       = element_line(color = "grey80", linewidth = 0.4),
    axis.ticks.length  = unit(3, "pt"),
    plot.title         = element_blank(),
    axis.title         = element_text(size = 15, color = "grey40"),
    axis.text          = element_text(size = 11,  color = "grey50")
  )
countplot_2
ggsave("countplot_2.png",countplot_2, width = 5, height = 4, dpi = 600)
```
## This does not seem alarming at all. Of cause some changes are expected, but visually, the change in amount of sentences does not seem abtrupt, which is what we expect, at the data shoudl originate from the same source.


# SOme more on distributional changes:

```{r}
library(dplyr)
library(zoo)
library(ggplot2)

monthly_data <- inf_data %>%
  mutate(
    month = as.yearmon(date),
    post_2019 = year(date) >= 2019
  ) %>%
  group_by(month) %>%
  summarise(n_sentences = n(), post_2019 = first(post_2019))
```


```{r}
density_plot <- ggplot(monthly_data, aes(x = n_sentences, fill = factor(post_2019))) +
  geom_density(alpha = 0.4) +
  labs(fill = "Post 2019") +
  xlab("Number of Sentences")+
  ylab("Density")+
  theme_minimal()+
  theme(
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = "grey92", linewidth = 0.4),
    axis.line.x        = element_line(color = "grey80", linewidth = 0.4),
    axis.ticks.x       = element_line(color = "grey80", linewidth = 0.4),
    axis.ticks.length  = unit(3, "pt"),
    plot.title         = element_blank(),
    axis.title         = element_text(size = 15, color = "grey40"),
    axis.text          = element_text(size = 13,  color = "grey50")
  )

density_plot
ggsave("density_plot.png",density_plot,width = 5, height = 4, dpi = 600)
```

```{r}
boxplot_plot <- ggplot(monthly_data, aes(x = factor(post_2019), y = n_sentences, fill = post_2019)) +
  geom_boxplot() +
  xlab("Post 2019")+
  ylab("Number of Sentences")+
  theme_minimal()+
   theme(
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = "grey92", linewidth = 0.4),
    axis.line.x        = element_line(color = "grey80", linewidth = 0.4),
    axis.ticks.x       = element_line(color = "grey80", linewidth = 0.4),
    axis.ticks.length  = unit(3, "pt"),
    plot.title         = element_blank(),
    axis.title         = element_text(size = 15, color = "grey40"),
    axis.text          = element_text(size = 13,  color = "grey50"),
    legend.position = "none"
  )

boxplot_plot
ggsave("boxplot.png",boxplot_plot,width = 5, height = 4, dpi = 600)
```
# Instead of comparing raw distributions, compare residuals after removing the trend:

```{r}
monthly_data <- monthly_data %>%
  mutate(t = 1:n())

trend_model <- lm(n_sentences ~ t, data = monthly_data)

monthly_data$residuals <- resid(trend_model)

ggplot(monthly_data, aes(x = residuals, fill = factor(post_2019))) +
  geom_density(alpha = 0.4)+ theme_minimal()
```
# and formally test

```{r}
ks.test(
  monthly_data$residuals[monthly_data$post_2019 == 0],
  monthly_data$residuals[monthly_data$post_2019 == 1]
)
```
# Can also be this approach:

```{r}
extra_model <- lm(n_sentences ~ t + post_2019, data = monthly_data)
summary(extra_model)

```

