---
title: "analysis_v2"
output: html_document
date: "2026-04-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#load packages

```{r, echo=FALSE, include=FALSE}
install.packages("pacman")
library(pacman)

pacman::p_load("tidyverse", "lubridate", "zoo", "glmmTMB", "DHARMa", "emmeans", "sjPlot", "modelsummary")

```


```{r}
#assumption check function
overdisp_fun <- function(model) {
  rdf <- df.residual(model)
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}


model_check <- function(model_name){
  hist(resid(model_name))
  
  overdisp_fun(model_name)
  
  plot(fitted(model_name), resid(model_name)) #residuals vs fitted
  abline(h=0)
  
  sim_res <- simulateResiduals(model_name, n = 1000)

  # QQ plot
  #plotQQunif(sim_res)   # QQ plot for simulated residuals
  plot(sim_res) 
  testUniformity(sim_res)      # Overall distribution
  testDispersion(sim_res)      # Over/underdispersion
  testZeroInflation(sim_res)   # Excess zeros
  testOutliers(sim_res, type = 'bootstrap')        # Outlying observations
  
  print(overdisp_fun(model_name))
  
}
```



#load data
```{r}
#inf_data = read_csv("/work/MarkusLundsfrydJensen#1865/Bachelor_project/data/inference/predicted_inference_data_with_gov.csv")
inf_data = read_csv("../../data/inference/predicted_inference_data_with_gov.csv")

#Make date column into date object
inf_data$date = as.Date(inf_data$date)

#Get the month
inf_data$month <- as.yearmon(inf_data$date)
# Extract only the year
inf_data$year <- as.numeric(format(inf_data$month, "%y"))

inf_data$party = as.factor(inf_data$party)
```

```{r}
head(inf_data_filtered)
```


```{r}

#Omitted parties
#NA, "UP", "FRI" "CD"  "LH"  "FP"  "IA"  "SIU" "T"   "UFG" "FF" "SP" "NQ"  "JF"   , "-"
parties_to_keep <- c("S", "DF", "EL", "SF", "V", "RV", "KD", "LA", "DD", "KF", "ALT", "NB", "M" )

#name parties that changed name
inf_data$party[inf_data$party == 'KRF'] <- 'KD'

#Ny alliance og LA
inf_data$party[inf_data$party == 'NY'] <- 'LA'

inf_data_filtered <- inf_data %>% 
  filter(party %in% parties_to_keep) %>% 
  mutate(party = droplevels(party))

```

```{r}
inf_data_filtered
```



#Make personal leaderboard
```{r}
blame_leaderboard <- inf_data_filtered %>% 
  group_by(speaker) %>% 
  summarise(total_blame = sum(prediction),
            n_sent = n(),
            blame_percentage = (total_blame/n_sent)*100,
            parties = list(unique(party))) %>% 
  filter(n_sent>=1000) %>% 
  ungroup() %>% 
  arrange(desc(blame_percentage))


head(blame_leaderboard, 10)
```

#Make party leaderboard

```{r}
party_leaderboard <- inf_data_filtered %>% 
  group_by(party) %>% 
  summarise(party_n_sent = n(),
            party_blame = sum(prediction),
            party_blame_percentage = (party_blame/party_n_sent)*100) %>% 
  filter(party_n_sent>=1000) %>% 
  arrange(desc(party_blame_percentage)) %>% 
  ungroup()

head(party_leaderboard)
```


```{r, echo=FALSE}
#make montly blame count pr party along with number of sentences

monthly_blame <- inf_data_filtered %>% 
  group_by(month, party) %>% 
  summarise(in_gov = first(in_gov),
            total_nr_sentences = n(),
            total_blame = sum(prediction),
            year = unique(year)) %>% 
  ungroup()
```


```{r}
# Read the extracted pairs
party_wingness <- read.csv("../python/bloc_and_wingness.csv")

party_wingness <- party_wingness %>% 
  select(party, bloc, wingness) %>% 
  mutate(
    bloc = as.character(bloc),  # avoid factor issues
    bloc = case_when(
      bloc == "left_bloc" ~ "right_bloc",
      bloc == "right_bloc" ~ "left_bloc",
      TRUE ~ bloc
    ),
    bloc = as.factor(bloc)
  )

table(party_wingness$bloc)


print(head(party_wingness))

print(head(monthly_blame))

# Merge/append to your existing R dataframe by party
monthly_blame <- merge(monthly_blame, party_wingness, by = "party", all.x = TRUE)

print(head(monthly_blame))
print(str(monthly_blame))
```

```{r}
monthly_blame$bloc <- as.factor(monthly_blame$bloc)

monthly_blame$in_gov <- as.factor(monthly_blame$in_gov)

monthly_blame$total_blame <- as.integer(monthly_blame$total_blame)

monthly_blame$wingness <- as.numeric(monthly_blame$wingness)

monthly_blame$year <- as.integer(monthly_blame$year)

monthly_blame$month_index <- with(monthly_blame,
  12 * (as.integer(format(month, "%Y")) - as.integer(format(min(month), "%Y"))) +
  (as.integer(format(month, "%m")) - as.integer(format(min(month), "%m")))
)

monthly_blame$month_index_scaled <- scale(monthly_blame$month_index)
monthly_blame$month_index_scaled_sqrt <- scale(monthly_blame$month_index^2)


```

```{r}
print(str(monthly_blame))
```





#Blame predicted by time


```{r}
model_blame_time_1 <- glmmTMB(total_blame ~ 1 + in_gov + (1|party) + offset(log(total_nr_sentences)), 
                           family = nbinom1, 
                           data = monthly_blame)
summary(model_blame_time_1)
diagnose(model_blame_time_1)
```



```{r}
model_blame_time_2 <- glmmTMB(total_blame ~ scale(month_index) + in_gov + (1|party) + offset(log(total_nr_sentences)), 
                           family = nbinom1, 
                           data = monthly_blame)
summary(model_blame_time_2)
diagnose(model_blame_time_2)
```

```{r}
model_blame_time_3 <- glmmTMB(total_blame ~ scale(month_index) + I(scale(month_index)^2) + in_gov + (1|party) + offset(log(total_nr_sentences)),
                           family = nbinom1, 
                           data = monthly_blame)

summary(model_blame_time_3)
diagnose(model_blame_time_3)
```



```{r}
anova(model_blame_time_1, model_blame_time_2, model_blame_time_3)
```

#check assumptions
```{r}
model_check(model_blame_time_3)
```

#Find global minimum

```{r}
b1 <- fixef(model_blame_time_3)$cond["scale(month_index)"]
b2 <- fixef(model_blame_time_3)$cond["I(scale(month_index)^2)"]

x_star <- -b1 / (2 * b2)

m <- mean(monthly_blame$month_index)
s <- sd(monthly_blame$month_index)

month_star <- x_star * s + m

print(month_star)

print(231/12)
```


```{r}
b1 <- fixef(model_blame_time_4)$cond["scale(month_index)"]
b2 <- fixef(model_blame_time_4)$cond["I(scale(month_index)^2)"]
b3 <- fixef(model_blame_time_4)$cond["I(scale(month_index)^3)"]

# derivative: b1 + 2*b2*x + 3*b3*x^2 = 0
# solve with quadratic formula
discriminant <- (2*b2)^2 - 4*(3*b3)*b1

if (discriminant < 0) {
  message("No real turning points")
} else {
  x1 <- (-2*b2 + sqrt(discriminant)) / (2 * 3*b3)
  x2 <- (-2*b2 - sqrt(discriminant)) / (2 * 3*b3)
  
  # back-transform from scaled to original month index
  m <- mean(monthly_blame$month_index)
  s <- sd(monthly_blame$month_index)
  
  month_star1 <- x1 * s + m
  month_star2 <- x2 * s + m
  
  print(month_star1)
  print(month_star2)
}
```




#regarding random effects:
The variance of the random slope is very small (0.009 vs 0.061 for the intercept), meaning parties barely differ in how they respond to government status — the random slope is adding very little while destabilising the model. Switched to simple intercept for party.
The random intercept alone is the defensible and stable choice here. You can justify dropping the random slope by noting that the variance component was negligible and the model showed signs of overparameterisation given 14 groups — this is a principled simplification, not a shortcut.

#regarding zero deflation:
Option B — Keep nbinom1 and acknowledge the limitation:
Simply note in your methods section that the model shows mild zero deflation affecting 1.2% of observations, that this does not materially affect your substantive conclusions, and that dispersion is otherwise well specified. Reviewers will accept this given the scale of the issue.

Practically Speaking
Given that dispersion is fine and the KS test is the only remaining concern, Option B is likely sufficient for most research contexts. The KS test is sensitive enough that 31 mishandled zeros out of 2538 observations can produce a significant result even when the model is otherwise well specified. This is a known limitation of the test at large sample sizes rather than evidence of a serious model misspecification.



```{r}
# Raw count
table(monthly_blame$total_blame == 0)

# As proportion
mean(monthly_blame$total_blame == 0)

# Which parties produce zeros?
monthly_blame %>% 
  group_by(party) %>% 
  summarise(zero_months = sum(total_blame == 0),
            total_months = n(),
            zero_pct = mean(total_blame == 0))
```


#regarding nbinom 1 or 2:
For parliamentary speech data, nbinom1 is often the better fit because blame counts tend to scale proportionally with how much a party speaks — which your offset already partially controls for. The fact that nbinom1 is already outperforming nbinom2 in your models is consistent with this.

#Plot best model

#OBS plot intercept with confidence intervals. Also change colros of random effects to match red/blue bloc, OBS BIas correction

```{r}
#install.packages("sjPlot")

library(sjPlot)

#Due to the offset, in order to do meaningfull interpretation, the total number of sentences must be kept constant. Thus the following plot shows the change in predicted amount of blame pr 1000 sentences uttered as time goes on  
sjPlot::plot_model(model_blame_time_3, 
                   type = "pred", 
                   terms = "month_index[all]", 
                   condition = c(total_nr_sentences = 100),
                   axis.title = c("Months post October 1997", "Predicted Blame Counts pr 100 Uttered Sentences"),
                   #title = "General Evolution of Blame in the Danish Parliament over time",
                   ci.lvl = .95,
                   #show.data = TRUE,
                   axis.lim = c(7,14)
                   ) +
  theme_minimal()
```
#try overlapping model with raw data

```{r}
plot_data <- monthly_blame %>%
  mutate(blame_rate = (total_blame / total_nr_sentences) * 100)

sjPlot::plot_model(model_blame_time_3,
                   type = "pred",
                   terms = "month_index[all]",
                   condition = c(total_nr_sentences = 100),
                   ci.lvl = .95) +
  geom_point(data = plot_data,
             aes(x = month_index, y = blame_rate),
             alpha = 0.2, size = 0.8, color = "steelblue") +
  theme_minimal()
```


```{r}
monthly_avg <- monthly_blame %>%
  group_by(month_index) %>%
  summarise(blame_rate = sum(total_blame) / sum(total_nr_sentences) * 100)

sjPlot::plot_model(model_blame_time_3,
                   type = "pred",
                   terms = "month_index[all]",
                   condition = c(total_nr_sentences = 100),
                   ci.lvl = .95) +
  geom_smooth(data = monthly_avg,
              aes(x = month_index, y = blame_rate),
              color = "steelblue", se = FALSE,
              linewidth = 0.7) +
  theme_minimal()
```

# Interpretation



#Now for political predictors of blame
Given your research question in Family 2 is about political characteristics, time is a confounder to control for, not a grouping structure to model. A fixed effect is simpler and more appropriate:


```{r}
model_intercept <- glmmTMB(total_blame ~ 1 + scale(month_index) + I(scale(month_index)^2) + (1|party) + offset(log(total_nr_sentences)), 
                           family = nbinom1, 
                           data = monthly_blame)

summary(model_intercept)
diagnose(model_intercept)
```

```{r}
model_in_gov <- glmmTMB(total_blame ~ in_gov + scale(month_index) + I(scale(month_index)^2) + (1 | party) +
                          offset(log(total_nr_sentences)),
                        family = nbinom1,
                        data = monthly_blame)

summary(model_in_gov)
diagnose(model_in_gov)
```



```{r}
model_bloc <- glmmTMB(total_blame ~ in_gov + bloc + scale(month_index) + I(scale(month_index)^2) + (1 | party) +
                         offset(log(total_nr_sentences)),
                       family = nbinom1,
                       data = monthly_blame)

summary(model_bloc)
diagnose(model_bloc)
```

```{r}
model_bloc_wing <- glmmTMB(total_blame ~ in_gov + bloc + wingness + scale(month_index) + I(scale(month_index)^2) + (1 | party) +
                         offset(log(total_nr_sentences)),
                       family = nbinom1,
                       data = monthly_blame)

summary(model_bloc_wing)
diagnose(model_bloc_wing)
```


```{r}
model_bloc_wing_int <- glmmTMB(total_blame ~ in_gov + bloc * wingness + scale(month_index) + I(scale(month_index)^2) + (1 | party) +
                             offset(log(total_nr_sentences)),
                           family = nbinom1,
                           data = monthly_blame)

summary(model_bloc_wing_int)
```

```{r}
model_bloc_bloc_wing_int <- glmmTMB(total_blame ~ in_gov*bloc + bloc * wingness + scale(month_index) + I(scale(month_index)^2) + (1 | party) +
                             offset(log(total_nr_sentences)),
                           family = nbinom1,
                           data = monthly_blame)

summary(model_bloc_bloc_wing_int)
```
```{r}
exp(0.553758)
```



```{r}
anova(model_intercept,model_in_gov,model_bloc,model_bloc_wing,model_bloc_wing_int, model_bloc_bloc_wing_int)
```

```{r}
anova(model_in_gov, model_bloc_wing, model_bloc_wing_int, model_bloc_bloc_wing_int)
```

```{r}
anova(model_in_gov, model_bloc_wing_int, model_bloc_bloc_wing_int)
```

#check assumptions of best model

```{r}
model_check(model_bloc_bloc_wing_int)
```


```{r}
summary(model_bloc_bloc_wing_int)
```



#plot it

```{r}
model_bloc_bloc_wing_int <- glmmTMB(total_blame ~ in_gov*bloc + bloc * wingness + month_index_scaled + month_index_scaled_sqrt + (1 | party) +
                             offset(log(total_nr_sentences)),
                           family = nbinom1,
                           data = monthly_blame)

summary(model_bloc_bloc_wing_int)
```



```{r}
sjPlot::plot_model(model_bloc_bloc_wing_int, 
                   type = "pred", 
                   terms = c("in_gov", "bloc"), 
                   condition = c(total_nr_sentences = 100),
                   axis.title = c("In government", "Predicted Blame Counts pr 100 Uttered Sentences"),
                   title = "Blame Predicted by Government status and ideological bloc",
                   ci.lvl = .95,
                   #axis.lim = c(3,12)
                   ) +
  theme_minimal()


sjPlot::plot_model(model_bloc_bloc_wing_int, 
                   type = "pred", 
                   terms = c("wingness"), 
                   condition = c(total_nr_sentences = 100),
                   axis.title = c("Wingness", "Predicted Blame Counts pr 100 Uttered Sentences"),
                   title = "Blame Predicted by Government status and ideological bloc",
                   ci.lvl = .95,
                   #axis.lim = c(3,12)
                   ) +
  theme_minimal()

sjPlot::plot_model(model_bloc_bloc_wing_int, 
                   type = "pred", 
                   terms = c("bloc","wingness"), 
                   condition = c(total_nr_sentences = 100),
                   axis.title = c("Bloc", "Predicted Blame Counts pr 100 Uttered Sentences"),
                   title = "Blame Predicted by Government status and ideological bloc",
                   
                   ci.lvl = .95,
                   #axis.lim = c(3,12)
                   ) +
  theme_minimal()
```



# DO check to see if quadratic control variable for time alters focal predicters


```{r}
# Compare linear vs quadratic time control in Family 2
model_block_int_linear <- glmmTMB(total_blame ~ in_gov*bloc + bloc * wingness + month_index + (1 | party) +
                             offset(log(total_nr_sentences)),
                           family = nbinom1,
                           data = monthly_blame)

model_block_int_quadratic <- glmmTMB(total_blame ~ in_gov*bloc + bloc * wingness + scale(month_index) + I(scale(month_index)^2) + (1 | party) +
                             offset(log(total_nr_sentences)),
                           family = nbinom1,
                           data = monthly_blame)

AIC(model_block_int_linear, model_block_int_quadratic)

anova(model_block_int_linear, model_block_int_quadratic)
```


```{r}
# Also check whether political coefficients change materially
summary(model_block_int_linear)

```


```{r}
summary(model_block_int_quadratic)
```



# For newer times


#Make data

```{r}
monthly_blame_18 <- monthly_blame %>% 
  filter(year >=19) %>% 
  filter(year <=26) %>% 
  mutate(month_index = month_index-255)
```




#Do analysis
monthly_blame_18
`
```{r}
model_intercept_18 <- glmmTMB(total_blame ~ 1 + month_index + (1|party)+ offset(log(total_nr_sentences)), 
                           family = nbinom1, 
                           data = monthly_blame_18)

summary(model_intercept_18)
```

```{r}
model_in_gov_18 <- glmmTMB(total_blame ~ in_gov + month_index + (1 | party) +
                          offset(log(total_nr_sentences)),
                        family = nbinom1,
                        data = monthly_blame_18)

summary(model_in_gov_18)
```



```{r}
model_bloc_18 <- glmmTMB(total_blame ~ in_gov + bloc + month_index + (1 | party) +
                         offset(log(total_nr_sentences)),
                       family = nbinom1,
                       data = monthly_blame_18)

summary(model_bloc_18)
```

```{r}
model_bloc_wing_18 <- glmmTMB(total_blame ~ in_gov + bloc + wingness + month_index + (1 | party) +
                         offset(log(total_nr_sentences)),
                       family = nbinom1,
                       data = monthly_blame_18)

summary(model_bloc_wing_18)
```

```{r}
model_bloc_wing_int_18 <- glmmTMB(total_blame ~ in_gov + bloc * wingness + month_index + (1 | party) +
                             offset(log(total_nr_sentences)),
                           family = nbinom1,
                           data = monthly_blame_18)

summary(model_bloc_wing_int_18)
```


```{r}
model_in_gov_bloc_wing_int_18 <- glmmTMB(total_blame ~ in_gov *bloc + bloc * wingness + month_index + (1 | party) +
                             offset(log(total_nr_sentences)),
                           family = nbinom1,
                           data = monthly_blame_18)

summary(model_in_gov_bloc_wing_int_18)
```




```{r}
anova(model_intercept_18, model_in_gov_18, model_bloc_18, model_bloc_wing_18, model_bloc_wing_int_18, model_in_gov_bloc_wing_int_18)
```

```{r}
anova(model_in_gov_18, model_bloc_wing_int_18, model_in_gov_bloc_wing_int_18)
```



```{r}
anova(model_in_gov_18, model_bloc_wing_int_18)
```

```{r}
#best model
summary(model_bloc_wing_int_18)
```
```{r}
exp(0.8171468)
```

#check assumption of best model

```{r}
model_check(model_bloc_wing_int_18)
```
#Plot best model

```{r}
sjPlot::plot_model(model_bloc_wing_int_18, 
                   type = "pred", 
                   terms = c("in_gov", "bloc"), 
                   condition = c(total_nr_sentences = 100),
                   axis.title = c("In Government", "Predicted Blame Counts pr 100 Uttered Sentences"),
                   title = "Predicted Blame by In Government and ideological Bloc in period 2019-2022",
                   ci.lvl = .95,
                   #axis.lim = c(3,12)
                   ) +
  theme_minimal()
sjPlot::plot_model(model_bloc_wing_int_18, 
                   type = "pred", 
                   terms = c("in_gov"), 
                   condition = c(total_nr_sentences = 100),
                   axis.title = c("In Government", "Predicted Blame Counts pr 100 Uttered Sentences"),
                   title = "Predicted Blame by In Government and ideological Bloc in period 2019-2022",
                   ci.lvl = .95,
                   #axis.lim = c(3,12)
                   ) +
  theme_minimal()

sjPlot::plot_model(model_bloc_wing_int_18, 
                   type = "pred", 
                   terms = c("bloc"), 
                   condition = c(total_nr_sentences = 100),
                   axis.title = c("Bloc", "Predicted Blame Counts pr 100 Uttered Sentences"),
                   title = "Predicted Blame by In Government and ideological Bloc in period 2019-2022",
                   ci.lvl = .95,
                   #axis.lim = c(3,12)
                   ) +
  theme_minimal()

sjPlot::plot_model(model_bloc_wing_int_18, 
                   type = "pred", 
                   terms = c("wingness"), 
                   condition = c(total_nr_sentences = 100),
                   axis.title = c("Wingness", "Predicted Blame Counts pr 100 Uttered Sentences"),
                   title = "Predicted Blame by Wing status in period 2019-2022",
                   ci.lvl = .95,
                   #axis.lim = c(3,12)
                   ) +
  theme_minimal()

sjPlot::plot_model(model_bloc_wing_int_18, 
                   type = "pred", 
                   terms = c("bloc","wingness"), 
                   condition = c(total_nr_sentences = 100),
                   axis.title = c("bloc", "Predicted Blame Counts pr 100 Uttered Sentences"),
                   title = "Predicted Blame by Wing status in period 2019-2022",
                   ci.lvl = .95,
                   #axis.lim = c(3,12)
                   ) +
  theme_minimal()
```




#time as predictor of blame in recent years

```{r}
model_blame_time_18_intercept <- glmmTMB(total_blame ~ 1 + in_gov  + (1 | party) + offset(log(total_nr_sentences)), 
                           family = nbinom1, 
                           data = monthly_blame_18)

summary(model_blame_time_18_intercept)
```

```{r}
model_blame_time_18 <- glmmTMB(total_blame ~  in_gov + scale(month_index)  + (1 | party) + offset(log(total_nr_sentences)), 
                           family = nbinom1, 
                           data = monthly_blame_18)

summary(model_blame_time_18)
```

```{r}
model_blame_time_18_2 <- glmmTMB(total_blame ~ in_gov + scale(month_index) + I(scale(month_index)^2)  + (1 | party) + offset(log(total_nr_sentences)), 
                           family = nbinom1, 
                           data = monthly_blame_18)

summary(model_blame_time_18_2)
```
```{r}
anova(model_blame_time_18_intercept,model_blame_time_18, model_blame_time_18_2)
```

```{r}
model_check(model_blame_time_18)
```

# OBS due to convergence issues/singular fit, nbinom 2 was needed here

```{r}
#best model
summary(model_blame_time_18)
```
```{r}
exp(0.092078)
```

#Plot this model

```{r}
sjPlot::plot_model(model_blame_time_18, 
                   type = "pred", 
                   terms = c("month_index"), 
                   condition = c(total_nr_sentences = 100),
                   axis.title = c("Months Post Jan 2019", "Predicted Blame Counts pr 100 Uttered Sentences"),
                   title = "Evolution of the Use of Blame in period 2019-2022",
                   ci.lvl = .95,
                    axis.lim = c(7,14)
                   ) +
  theme_minimal()
```

```{r}
sum(monthly_blame$total_blame)/sum(monthly_blame$total_nr_sentences)
```


```{r}
table(monthly_blame$bloc, monthly_blame$wingness)
```

#get model summary in latex format

```{r}
summary(model_blame_time_3)
```




```{r}
modelsummary(
  model_blame_time_3,
  fmt = fmt_significant(3),          # 3 significant digits
  exponentiate = FALSE,              # set TRUE if you want IRRs
  statistic = "std.error",           # show SE (or "statistic", "p.value", "conf.int")
  stars = TRUE,
  output = "latex",
  title = "Negative Binomial Model Predicting Total Blame (1997-2026)",
  notes = "Random effect: party (intercept). Offset: log(total\\_nr\\_sentences)."
)

# Save directly to file:
modelsummary(
  model_blame_time_3,
  fmt = fmt_significant(3),
  output = "analysis_model_time_summary.tex"
)
```









```{r}

modelsummary(
  model_blame_time_3,
  output = "latex"
)
```

#get model summary in latex format:
```{r}

modelsummary(
  model_bloc_bloc_wing_int,
  output = "latex"
)
```

#get model summary in latex format:
```{r}

modelsummary(
  model_blame_time_18,
  output = "latex"
)
```

#get model summary in latex format:
```{r}

modelsummary(
  model_bloc_wing_int_18,
  output = "latex"
)
```

#all latex models


```{r}
pacman::p_load("texreg")

texreg(model_blame_time_3, digits = 10, file = "analysis_model_time_summary.tex")
texreg(model_bloc_bloc_wing_int, digits = 10, file = "analysis_model_pol_summary.tex")
texreg(model_blame_time_18, digits = 10, file = "analysis_model_time_19_summary.tex")
texreg(model_bloc_wing_int_18, digits = 10, file = "analysis_model_pol_19_summary.tex")
```

```{r}
pacman::p_load("stringr", "texreg")

# Step 1: capture texreg output with maximum digits
tex_lines <- capture.output(
  texreg(model_blame_time_3, digits = 10)
)

# Step 2: handle coefficient lines with stars: $-2.3806^{***}$
# → \num{-2.3806}$^{***}$
tex_lines_fixed <- str_replace_all(
  tex_lines,
  pattern     = "\\$(-?\\d+\\.\\d+)(\\^\\{\\*+\\})\\$",
  replacement = "\\\\num{\\1}$\\2$"
)

# Step 3: handle plain number in math mode (no stars): $-2.3806$
# → \num{-2.3806}
tex_lines_fixed <- str_replace_all(
  tex_lines_fixed,
  pattern     = "\\$(-?\\d+\\.\\d+)\\$",
  replacement = "\\\\num{\\1}"
)

# Step 4: handle standard errors in parentheses: $(0.0644)$
# → (\num{0.0644})
tex_lines_fixed <- str_replace_all(
  tex_lines_fixed,
  pattern     = "\\$\\((-?\\d+\\.\\d+)\\)\\$",
  replacement = "(\\\\num{\\1})"
)

# Step 5: fix footnote line — restore it to original clean format
# target only the \scriptsize footnote line and rewrite it entirely
tex_lines_fixed <- str_replace_all(
  tex_lines_fixed,
  pattern     = "\\\\scriptsize\\{.*\\}",
  replacement = "\\\\scriptsize{$^{***}$p$<$0.001; $^{**}$p$<$0.01; $^{*}$p$<$0.05}"
)

# Step 6: save and inspect
cat(paste(tex_lines_fixed, collapse = "\n"))
writeLines(tex_lines_fixed, "test.tex")
```


# COmbining plots:

```{r}
library(sjPlot)
library(ggplot2)
library(patchwork)

p1 <- sjPlot::plot_model(
  model_blame_time_3,
  type = "pred",
  terms = "month_index[all]",
  condition = c(total_nr_sentences = 100),
  ci.lvl = .95,
  axis.lim = c(7, 14)
) +
  theme_minimal() +
  theme(
    axis.title.y = element_text(size = 10),
    axis.title.x = element_blank(),
    #plot.title   = element_blank()
  ) +
  labs(title = "Full Period (1997–2026)",
       y = "Blame Count per 100 Uttered Sentences")

p2 <- sjPlot::plot_model(
  model_blame_time_18,
  type = "pred",
  terms = "month_index[all]",
  condition = c(total_nr_sentences = 100),
  ci.lvl = .95,
  axis.lim = c(7, 14)
) +
  theme_minimal() +
  theme(
    axis.title.y  = element_blank(),
    axis.text.y   = element_blank(),
    axis.ticks.y  = element_blank(),
    axis.title.x  = element_blank(),
    #plot.title    = element_blank()
  ) +
  labs(title = "Recent Period (2019–2026)")

combined <- p1 + p2 +
  plot_layout(ncol = 2) +
  plot_annotation(
    caption = "Months post October 1997"
  ) &
  theme(
    plot.caption = element_text(hjust = 0.5, size = 10)
  )

combined
```


```{r}
library(sjPlot)
library(ggplot2)
library(patchwork)

separator <- ggplot() +
  geom_vline(xintercept = 0.5, linewidth = 0.3, colour = "grey60") +
  theme_void()

my_theme <- theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = "grey92", linewidth = 0.4),
    axis.line.x        = element_line(color = "grey80", linewidth = 0.4),
    axis.ticks.x       = element_line(color = "grey80", linewidth = 0.4),
    axis.ticks.length  = unit(3, "pt"),
    plot.title         = element_blank(),
    axis.title         = element_text(size = 15, color = "grey40"),
    axis.text          = element_text(size = 13,  color = "grey50")
  )

p1 <- sjPlot::plot_model(
  model_blame_time_3,
  type      = "pred",
  terms     = "month_index[all]",
  condition = c(total_nr_sentences = 100),
  ci.lvl    = .95,
  axis.lim  = c(7, 14),
  colors    = "#2C7BB6"
) +
  my_theme +
  theme(
    axis.title.x = element_blank()
  ) +
  labs(y     = "Predicted Blame (%)"
  )

p2 <- sjPlot::plot_model(
  model_blame_time_18,
  type      = "pred",
  terms     = "month_index[all]",
  condition = c(total_nr_sentences = 100),
  ci.lvl    = .95,
  axis.lim  = c(7, 14),
  colors    = "#2C7BB6"
) +
  my_theme +
  theme(
    axis.title.y  = element_blank(),
    axis.text.y   = element_blank(),
    axis.ticks.y  = element_blank(),
    axis.title.x  = element_blank()
  ) 

combined <- (p1 + separator + p2) +
  plot_layout(guides = "collect",
    widths = c(1, 0.02, 1)
  ) +
  plot_annotation(
    caption = "Months post October 1997"
  ) &
  theme(
    plot.caption = element_text(hjust = 0.5, size = 14, color = "grey40")
  )


ggsave("time_plots.png", combined, width = 5, height = 4)
combined
```


# COmbined plot for political characteristics

```{r}
# # Rename bloc levels before plotting
# monthly_blame$bloc <- factor(monthly_blame$bloc,
#                          levels = c("left_bloc", "right_bloc"),
#                          labels = c("Left bloc", "Right bloc"))
# # Rename bloc levels before plotting
# monthly_blame_18$bloc <- factor(monthly_blame_18$bloc,
#                          levels = c("left_bloc", "right_bloc"),
#                          labels = c("Left bloc", "Right bloc"))
```





```{r}
ggsave("time_plots.png", combined, width = 5, height = 4)
```

```{r}
library(sjPlot)
library(ggplot2)
library(patchwork)

my_theme <- theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = "grey92", linewidth = 0.4),
    axis.line.x        = element_line(color = "grey80", linewidth = 0.4),
    axis.ticks.x       = element_line(color = "grey80", linewidth = 0.4),
    axis.ticks.length  = unit(3, "pt"),
    plot.title         = element_blank(),
    axis.title         = element_text(size = 15, color = "grey40"),
    axis.text          = element_text(size = 11, color = "grey50"),
    legend.position    = "bottom",
    legend.title       = element_text(size = 13, color = "grey40"),
    legend.text        = element_text(size = 13, color = "grey50")
  )

separator <- ggplot() +
  geom_vline(xintercept = 0.5, linewidth = 0.3, colour = "grey60") +
  theme_void()

wing_colors <- c("left_bloc"  = "#C0392B",
                 "right_bloc" = "#2C7BB6")

wingness_colors <- c("#9ECAE1",
                     "#2C7BB6",
                     "#08306B")

# ── Figure 1: gov * bloc, both periods ────────────────────────────────────────

p1 <- sjPlot::plot_model(
  model_bloc_bloc_wing_int,
  type      = "pred",
  terms     = c("in_gov", "bloc"),
  condition = c(total_nr_sentences = 100),
  ci.lvl    = .95,
  colors    = wing_colors,
  axis.lim  = c(4, 13)
) +
  my_theme +
  theme(axis.title.x = element_blank()) +
  labs(
    y     = "Predicted Blame (%)",
    color = "Wing",
    fill  = "Wing"
  ) +
  scale_color_manual(name   = "Wing Affiliation",
                     values = wing_colors,
                     labels = c("Left Wing", "Right Wing")) +
  scale_fill_manual(name    = "Wing Affiliation",
                    values  = wing_colors,
                    labels  = c("Left Wing", "Right Wing")) +
  scale_x_continuous(breaks = c(0, 1),
                     labels = c("Opposition", "Government"),
                     limits = c(-0.5, 1.5),
                     expand = c(0, 0))

p2 <- sjPlot::plot_model(
  model_bloc_wing_int_18,
  type      = "pred",
  terms     = c("in_gov", "bloc"),
  condition = c(total_nr_sentences = 100),
  ci.lvl    = .95,
  colors    = wing_colors,
  axis.lim  = c(4, 13)
) +
  my_theme +
  theme(
    axis.title.y = element_blank(),
    axis.text.y  = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.x = element_blank()
  ) +
  labs(
    color = "Wing",
    fill  = "Wing"
  ) +
  scale_color_manual(name   = "Wing Affiliation",
                     values = wing_colors,
                     labels = c("Left Wing", "Right Wing")) +
  scale_fill_manual(name    = "Wing Affiliation",
                    values  = wing_colors,
                    labels  = c("Left Wing", "Right Wing")) +
  scale_x_continuous(breaks = c(0, 1),
                     labels = c("Opposition", "Government"),
                     limits = c(-0.5, 1.5),
                     expand = c(0, 0))

# figure_1 <- (p1 + p2) +
#   plot_layout(guides = "collect") +
#   plot_annotation(caption = "Government Status") &
#   theme(
#     plot.caption    = element_text(hjust = 0.5, size = 14, color = "grey40"),
#     legend.position = "bottom"
#   )
p1$layers[[1]]$position <- position_dodge(width = 0.5)
p1$layers[[2]]$position <- position_dodge(width = 0.5)
p2$layers[[1]]$position <- position_dodge(width = 0.5)
p2$layers[[2]]$position <- position_dodge(width = 0.5)

figure_1 <- (p1 + separator + p2) +
  plot_layout(
    guides = "collect",
    widths = c(1, 0.02, 1)
  ) +
  plot_annotation(caption = "Government Status") &
  theme(
    plot.caption    = element_text(hjust = 0.5, size = 14, color = "grey40"),
    legend.position = "bottom"
  )

# ── Figure 2: bloc * wingness, both periods ───────────────────────────────────

p3 <- sjPlot::plot_model(
  model_bloc_bloc_wing_int,
  type      = "pred",
  terms     = c("bloc", "wingness [0.0, 0.75, 1.5]"),
  condition = c(total_nr_sentences = 100),
  ci.lvl    = .95,
  colors    = wingness_colors,
  axis.lim  = c(4, 22)
) +
  my_theme +
  theme(axis.title.x = element_blank()) +
  labs(
    y     = "Predicted Blame (%)",
    color = "Wingness",
    fill  = "Wingness"
  ) +
  scale_color_manual(name   = "Wingness",
                     values = wingness_colors,
                     labels = c("Low (0.0)", "Medium (0.75)", "High (1.5)")) +
  scale_fill_manual(name    = "Wingness",
                    values  = wingness_colors,
                    labels  = c("Low (0.0)", "Medium (0.75)", "High (1.5)")) +
  scale_x_continuous(breaks = c(1, 2),
                     labels = c("Left Wing", "Right Wing"),
                     limits = c(0.5, 2.5),
                     expand = c(0, 0))

p4 <- sjPlot::plot_model(
  model_bloc_wing_int_18,
  type      = "pred",
  terms     = c("bloc", "wingness [0.0, 0.75, 1.5]"),
  condition = c(total_nr_sentences = 100),
  ci.lvl    = .95,
  colors    = wingness_colors,
  axis.lim  = c(4, 22)
) +
  my_theme +
  theme(
    axis.title.y = element_blank(),
    axis.text.y  = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.x = element_blank()
  ) +
  labs(
    color = "Wingness",
    fill  = "Wingness"
  ) +
  scale_color_manual(name   = "Wingness",
                     values = wingness_colors,
                     labels = c("Low (0.0)", "Medium (0.75)", "High (1.5)")) +
  scale_fill_manual(name    = "Wingness",
                    values  = wingness_colors,
                    labels  = c("Low (0.0)", "Medium (0.75)", "High (1.5)")) +
  scale_x_continuous(breaks = c(1, 2),
                     labels = c("Left Wing", "Right Wing"),
                     limits = c(0.5, 2.5),
                     expand = c(0, 0))

# figure_2 <- (p3 + p4) +
#   plot_layout(guides = "collect") +
#   plot_annotation(caption = "Wing Affiliation") &
#   theme(
#     plot.caption    = element_text(hjust = 0.5, size = 14, color = "grey40"),
#     legend.position = "bottom"
#   )

p3$layers[[1]]$position <- position_dodge(width = 0.5)
p3$layers[[2]]$position <- position_dodge(width = 0.5)
p4$layers[[1]]$position <- position_dodge(width = 0.5)
p4$layers[[2]]$position <- position_dodge(width = 0.5)

figure_2 <- (p3 + separator + p4) +
  plot_layout(
    guides = "collect",
    widths = c(1, 0.02, 1)
  ) +
  plot_annotation(caption = "Wing Affiliation") &
  theme(
    plot.caption    = element_text(hjust = 0.5, size = 14, color = "grey40"),
    legend.position = "bottom"
  )

figure_1
figure_2
ggsave("gov_wing.png", figure_1, width = 5, height = 5)
ggsave("wing_wingness.png", figure_2, width = 5, height = 5)
```

```{r}
p3$layers[[1]]$position
```


