---
title: "analysis_v2"
output: html_document
date: "2026-04-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#load packages

```{r, echo=FALSE, include=FALSE}
install.packages("pacman")

library(pacman)

pacman::p_load("tidyverse", "lubridate", "zoo", "glmmTMB", "DHARMa", "emmeans", "sjPlot")

```


```{r}
#assumption check function
overdisp_fun <- function(model) {
  rdf <- df.residual(model)
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}


model_check <- function(model_name){
  hist(resid(model_name))
  
  overdisp_fun(model_name)
  
  plot(fitted(model_name), resid(model_name)) #residuals vs fitted
  abline(h=0)
  
  sim_res <- simulateResiduals(model_name, n = 1000)

  # QQ plot
  #plotQQunif(sim_res)   # QQ plot for simulated residuals
  plot(sim_res) 
  testUniformity(sim_res)      # Overall distribution
  testDispersion(sim_res)      # Over/underdispersion
  testZeroInflation(sim_res)   # Excess zeros
  testOutliers(sim_res, type = 'bootstrap')        # Outlying observations
  
  print(overdisp_fun(model_name))
  
}
```



#load data
```{r}
inf_data = read_csv("C:/Uni/Project/BlameBERT/data/inference/predicted_inference_data_with_gov.csv")

#Make date column into date object
inf_data$date = as.Date(inf_data$date)

#Get the month
inf_data$month <- as.yearmon(inf_data$date)
# Extract only the year
inf_data$year <- as.numeric(format(inf_data$month, "%y"))

inf_data$party = as.factor(inf_data$party)
```



```{r}

#Omitted parties
#NA, "UP", "FRI" "CD"  "LH"  "FP"  "IA"  "SIU" "T"   "UFG" "FF" "SP" "NQ"  "JF"   , "-"
parties_to_keep <- c("S", "DF", "EL", "SF", "V", "RV", "KD", "LA", "DD", "KF", "ALT", "NB", "M" )

#name parties that changed name
inf_data$party[inf_data$party == 'KRF'] <- 'KD'

#Ny alliance og LA
inf_data$party[inf_data$party == 'NY'] <- 'LA'

inf_data_filtered <- inf_data %>% 
  filter(party %in% parties_to_keep) %>% 
  mutate(party = droplevels(party))

inf_data_filtered <- inf_data_filtered %>%
  mutate(
    prediction_strict = as.integer(prediction == 1 & confidence > 0.7)  # raise from 0.5
  )

```


```{r}
strict <- inf_data_filtered %>% 
  
  prediction==1)
```


#Make personal leaderboard

## REVISIT WITH EM AFTER CHECKING IT WORKS
```{r}
blame_leaderboard <- inf_data_filtered %>% 
  group_by(speaker) %>% 
  summarise(total_blame = sum(prediction),
            n_sent = n(),
            blame_percentage = (total_blame/n_sent)*100,
            parties = list(unique(party))) %>% 
  filter(n_sent>=1000) %>% 
  ungroup() %>% 
  arrange(desc(blame_percentage))


head(blame_leaderboard, 10)
```

#Make party leaderboard

```{r}
party_leaderboard <- inf_data_filtered %>% 
  group_by(party) %>% 
  summarise(party_n_sent = n(),
            party_blame = sum(prediction),
            party_blame_percentage = (party_blame/party_n_sent)*100) %>% 
  filter(party_n_sent>=1000) %>% 
  arrange(desc(party_blame_percentage)) %>% 
  ungroup()

head(party_leaderboard)
```


```{r, echo=FALSE}
#make montly blame count pr party along with number of sentences

#NEW APPROACH - HIGHER CONFIDENCE SCORE REQUIRED

monthly_blame <- inf_data_filtered %>% 
  group_by(month, party) %>% 
  summarise(in_gov = first(in_gov),
            total_nr_sentences = n(),
            total_blame = sum(prediction_strict),
            year = unique(year)) %>% 
  ungroup()
```

```{r}
# RECALL = SENS
# PRECISION = SPEC
precision <- 0.72 # avg: 0.80
recall <- 0.79 # avg: 0.81 

monthly_blame <- monthly_blame %>%
  mutate(
    # Observed blame rate
    obs_blame_rate = total_blame / total_nr_sentences,
    
    # Corrected true blame rate using equation 1 from Magder & Hughes
    # p_true = (p_observed - (1 - precision)) / (recall - (1 - precision))
    true_blame_rate = (obs_blame_rate - (1 - precision)) / (recall - (1 - precision)),
    
    # Clamp to [0, 1] — if observed rate is below (1-spec), true rate is 0
    true_blame_rate = pmax(0.01, pmin(1, true_blame_rate)),
    
    # Reconstruct corrected count (keep as numeric, models handle this fine)
    total_blame_corrected = round(true_blame_rate * total_nr_sentences)
  )
```

```{r}
inf_data_filtered %>% slice(694)
```


```{r}
# NEW CHUNK which shows how different precision / recall scores would present  


# how many times more blame does a party produce per sentence when in government compared to when in opposition?
# So concretely:
# 
# IRR = 1.0 means no difference in blame rate between government and opposition
# IRR = 0.8 means parties in government produce blame at 80% the rate of opposition parties — i.e. 20% less blame per sentence
# IRR = 1.3 means parties in government produce 30% more blame per sentence than opposition parties

#INTUITION: The IRR (Incidence Rate Ratio) is a measure of how stable the correction math is essentially. It only really gives quantifiable measurements at a higher recall / precision. The highest row shows us that parties in government blame at about 0.5 the rate of opposition parties, which does align with our findings.


recall_vals <- c(0.75, 0.80, 0.85, 0.90, 0.95, 0.98)
precision_vals <- c(0.75, 0.80, 0.85, 0.90, 0.95, 0.98)

sensitivity_grid <- expand.grid(recall = recall_vals, precision = precision_vals) %>%
  rowwise() %>%
  mutate(
    coef_in_gov = {
      d <- monthly_blame %>%
        mutate(
          obs_rate = total_blame / total_nr_sentences,
          true_rate = pmax(0, (obs_rate - (1 - precision)) / (recall - (1 - precision))),
          total_blame_corrected = true_rate * total_nr_sentences
        )
      m <- glmmTMB(total_blame_corrected ~ in_gov*bloc + bloc*wingness +
                     month_index_scaled + month_index_scaled_sqrt + (1|party) +
                     offset(log(total_nr_sentences)),
                   family = nbinom1, data = d)
      fixef(m)$cond["in_gov1"]
    }
  ) %>%
  ungroup()

# Plot it
ggplot(sensitivity_grid, aes(x = recall, y = precision, fill = exp(coef_in_gov))) +
  geom_tile() +
  geom_text(aes(label = round(exp(coef_in_gov), 3)), size = 3, color="white") +
  scale_fill_viridis_c(name = "IRR for in_gov", option="h") +
  labs(title = "Sensitivity of in_gov effect to classifier performance assumptions",
       x = "Recall", y = "Precision") +
  theme_minimal()
```

```{r}
# You can see this threshold directly
tibble(precision = c(0.75, 0.80, 0.85, 0.90, 0.95)) %>%
  mutate(
    false_positive_rate = 1 - precision,
    expected_fp_blame_rate = 0.92 * false_positive_rate,
    correction_possible = expected_fp_blame_rate < 0.08  # your observed rate
  )

# Results will show:
# precision 0.75 -> expected FP rate 0.23 -> correction impossible
# precision 0.80 -> expected FP rate 0.18 -> correction impossible  
# precision 0.85 -> expected FP rate 0.14 -> correction impossible
# precision 0.90 -> expected FP rate 0.09 -> borderline
# precision 0.95 -> expected FP rate 0.05 -> correction possible
```


```{r}
# What is your overall observed blame rate?
mean(monthly_blame$obs_blame_rate)

# How often does correction go negative before clamping?
monthly_blame %>%
  mutate(
    obs_blame_rate = total_blame / total_nr_sentences,
    unclamped = (obs_blame_rate - (1 - precision)) / (recall - (1 - precision))
  ) %>%
  summarise(
    pct_negative = mean(unclamped < 0),
    pct_above_1  = mean(unclamped > 1),
    min_val      = min(unclamped),
    mean_val     = mean(unclamped)
  )
```

```{r}
# Just show what happens to your key IRRs under different assumptions
# rather than fully correcting the dependent variable

tibble(
  sens = c(0.70, 0.79, 0.80, 0.85, 0.90),
  spec = c(0.70, 0.72, 0.80, 0.85, 0.90)
) %>%
  mutate(
    bias_factor = 1 / (sens + spec - 1),
    note = "Your true ORs are approximately this multiple of your naive estimates"
  )
```

```{r}
# Quick sanity check with your actual numbers
sens <- 0.79
spec <- 0.72
true_rate <- 0.08

# What observed rate does your classifier actually produce?
expected_observed <- true_rate * sens + (1 - true_rate) * (1 - spec)
cat("Expected observed blame rate:", round(expected_observed, 3))
# = 0.08 * 0.79 + 0.92 * 0.28 = 0.063 + 0.258 = ~0.32
```





```{r}
# Read the extracted pairs
party_wingness <- read.csv("../python/bloc_and_wingness.csv")

party_wingness <- party_wingness %>% 
  select(party, bloc, wingness) %>% 
  mutate(
    bloc = as.character(bloc),  # avoid factor issues
    bloc = case_when(
      bloc == "left_bloc" ~ "right_bloc",
      bloc == "right_bloc" ~ "left_bloc",
      TRUE ~ bloc
    ),
    bloc = as.factor(bloc)
  )

table(party_wingness$bloc)


print(head(party_wingness))

print(head(monthly_blame))

# Merge/append to your existing R dataframe by party
monthly_blame <- merge(monthly_blame, party_wingness, by = "party", all.x = TRUE)

print(head(monthly_blame))
print(str(monthly_blame))
```

```{r}
monthly_blame$bloc <- as.factor(monthly_blame$bloc)

monthly_blame$in_gov <- as.factor(monthly_blame$in_gov)

monthly_blame$total_blame <- as.integer(monthly_blame$total_blame)

monthly_blame$total_blame_corrected <- as.integer(monthly_blame$total_blame_corrected)

monthly_blame$wingness <- as.numeric(monthly_blame$wingness)

monthly_blame$year <- as.integer(monthly_blame$year)

monthly_blame$month_index <- with(monthly_blame,
  12 * (as.integer(format(month, "%Y")) - as.integer(format(min(month), "%Y"))) +
  (as.integer(format(month, "%m")) - as.integer(format(min(month), "%m")))
)

monthly_blame$month_index_scaled <- scale(monthly_blame$month_index)
monthly_blame$month_index_scaled_sqrt <- scale(monthly_blame$month_index^2)


```

```{r}
print(str(monthly_blame))
```


```{r}
table(monthly_blame$total_blame_corrected)
summary(monthly_blame$total_blame_corrected)

length(unique(monthly_blame$party))  # confirm
table(monthly_blame$party)           # check if some parties have very few rows 
```

```{r}
# Are total_nr_sentences very large numbers?
summary(monthly_blame$total_nr_sentences)

# What is the actual blame rate?
mean(monthly_blame$total_blame_corrected / monthly_blame$total_nr_sentences)
```


#Blame predicted by time


```{r}
model_blame_time_1 <- glmmTMB(total_blame_corrected ~ 1 + in_gov + (1|party) + offset(log(total_nr_sentences)), 
                           family = "poisson",
                           data = monthly_blame)
summary(model_blame_time_1)
diagnose(model_blame_time_1)
```

```{r}
model_check(model_blame_time_1)
check_overdispersion(model_blame_time_1)

hist(monthly_blame$total_nr_sentences)
hist(log(monthly_blame$total_nr_sentences))
cor(monthly_blame$total_blame_corrected, monthly_blame$total_nr_sentences)
```

```{r}
monthly_blame$blame_rate <- monthly_blame$total_blame_corrected / monthly_blame$total_nr_sentences
hist(monthly_blame$blame_rate)
summary(monthly_blame$blame_rate)
```


```{r}
model_blame_time_2 <- glmmTMB(total_blame_corrected ~ scale(month_index) + in_gov + (1|party) + offset(log(total_nr_sentences)), 
                           family = "poisson",#nbinom1 
                           data = monthly_blame)
summary(model_blame_time_2)
diagnose(model_blame_time_2)
```

```{r}
model_blame_time_3 <- glmmTMB(total_blame_corrected ~ scale(month_index) + I(scale(month_index)^2) + in_gov + (1|party) + offset(log(total_nr_sentences)),
                           family = "poisson", #nbinom1
                           data = monthly_blame)

summary(model_blame_time_3)
diagnose(model_blame_time_3)
```



```{r}
anova(model_blame_time_1, model_blame_time_2, model_blame_time_3)
```
```{r}
# Distribution check
summary(monthly_blame$total_blame_corrected)

hist(monthly_blame$total_blame_corrected, breaks = 50,
     main = "Distribution of corrected blame counts")

# Compare variance to mean — should be >> 1 for nbinom to help
var(monthly_blame$total_blame_corrected) / mean(monthly_blame$total_blame_corrected)

# Check for non-integer values which can cause issues
head(monthly_blame$total_blame_corrected)
sum(monthly_blame$total_blame_corrected != round(monthly_blame$total_blame_corrected))

monthly_blame <- monthly_blame %>%
  mutate(total_blame_corrected = round(total_blame_corrected))
```

```{r}
# Checking for outliers in the corrected data
monthly_blame %>%
  arrange(desc(total_blame_corrected)) %>%
  select(party, month, total_blame, total_blame_corrected, total_nr_sentences) %>%
  head(40)
```

```{r}
# Compare variance-to-mean ratio before and after correction
monthly_blame %>%
  summarise(
    raw_vmr      = var(total_blame) / mean(total_blame),
    corrected_vmr = var(total_blame_corrected) / mean(total_blame_corrected)
  )
# Corrected VMR will be substantially higher
```

```{r}
monthly_blame %>%
  summarise(
    raw_vmr       = var(total_blame) / mean(total_blame),
    corrected_vmr = var(total_blame_corrected) / mean(total_blame_corrected),
    corrected_mean = mean(total_blame_corrected),
    corrected_var  = var(total_blame_corrected)
  )

# Also look at the actual shape
hist(monthly_blame$total_blame_corrected, breaks = 50)
hist(monthly_blame$total_blame, breaks = 50)
```




#check assumptions

```{r}
model_check(model_blame_time_3)
```

#Find global minimum

```{r}
b1 <- fixef(model_blame_time_3)$cond["scale(month_index)"]
b2 <- fixef(model_blame_time_3)$cond["I(scale(month_index)^2)"]

x_star <- -b1 / (2 * b2)

m <- mean(monthly_blame$month_index)
s <- sd(monthly_blame$month_index)

month_star <- x_star * s + m

print(month_star)

print(231/12)
```




#regarding random effects:
The variance of the random slope is very small (0.009 vs 0.061 for the intercept), meaning parties barely differ in how they respond to government status — the random slope is adding very little while destabilising the model. Switched to simple intercept for party.
The random intercept alone is the defensible and stable choice here. You can justify dropping the random slope by noting that the variance component was negligible and the model showed signs of overparameterisation given 14 groups — this is a principled simplification, not a shortcut.

#regarding zero deflation:
Option B — Keep nbinom1 and acknowledge the limitation:
Simply note in your methods section that the model shows mild zero deflation affecting 1.2% of observations, that this does not materially affect your substantive conclusions, and that dispersion is otherwise well specified. Reviewers will accept this given the scale of the issue.

Practically Speaking
Given that dispersion is fine and the KS test is the only remaining concern, Option B is likely sufficient for most research contexts. The KS test is sensitive enough that 31 mishandled zeros out of 2538 observations can produce a significant result even when the model is otherwise well specified. This is a known limitation of the test at large sample sizes rather than evidence of a serious model misspecification.



```{r}
# Raw count
table(monthly_blame$total_blame == 0)
#corrected
table(monthly_blame$total_blame_corrected==0)

# As proportion
mean(monthly_blame$total_blame == 0)
#corrected 
mean(monthly_blame$total_blame_corrected==0)

# Which parties produce zeros?
monthly_blame %>% 
  group_by(party) %>% 
  summarise(zero_months = sum(total_blame == 0),
            total_months = n(),
            zero_pct = mean(total_blame == 0))

#corrected
monthly_blame %>% 
  group_by(party) %>% 
  summarise(zero_months = sum(total_blame_corrected==0),
            total_months=n(),
            zero_pct=mean(total_blame_corrected==0))
```


#regarding nbinom 1 or 2:
For parliamentary speech data, nbinom1 is often the better fit because blame counts tend to scale proportionally with how much a party speaks — which your offset already partially controls for. The fact that nbinom1 is already outperforming nbinom2 in your models is consistent with this.

#Plot best model

#OBS plot intercept with confidence intervals. Also change colros of random effects to match red/blue bloc, OBS BIas correction

```{r}
#install.packages("sjPlot")

library(sjPlot)

sjPlot::plot_model(model_blame_time_3, type = "re")

#Due to the offset, in order to do meaningfull interpretation, the total number of sentences must be kept constant. Thus the following plot shows the change in predicted amount of blame pr 1000 sentences uttered as time goes on  
sjPlot::plot_model(model_blame_time_3, 
                   type = "pred", 
                   terms = "month_index[all]", 
                   condition = c(total_nr_sentences = 100),
                   axis.title = c("Months after Jan 2000", "Predicted Blame Counts pr 100 Uttered Sentences"),
                   title = "General Evolution of Blame in the Danish Parliament over time",
                   #ci.lvl = NA,
                   #axis.lim = c(3,12)
                   ) +
  theme_minimal()
```


# Interpretation



#Now for political predictors of blame
Given your research question in Family 2 is about political characteristics, time is a confounder to control for, not a grouping structure to model. A fixed effect is simpler and more appropriate:


```{r}
model_intercept <- glmmTMB(total_blame_corrected ~ 1 + scale(month_index) + I(scale(month_index)^2) + (1|party) + offset(log(total_nr_sentences)), 
                           family = nbinom1, 
                           data = monthly_blame)

summary(model_intercept)
diagnose(model_intercept)
```

```{r}
model_in_gov <- glmmTMB(total_blame_corrected ~ in_gov + scale(month_index) + I(scale(month_index)^2) + (1 | party) +
                          offset(log(total_nr_sentences)),
                        family = nbinom1,
                        data = monthly_blame)

summary(model_in_gov)
diagnose(model_in_gov)
```



```{r}
model_bloc <- glmmTMB(total_blame_corrected ~ in_gov + bloc + scale(month_index) + I(scale(month_index)^2) + (1 | party) +
                         offset(log(total_nr_sentences)),
                       family = nbinom1,
                       data = monthly_blame)

summary(model_bloc)
diagnose(model_bloc)
```

```{r}
model_bloc_wing <- glmmTMB(total_blame_corrected ~ in_gov + bloc + wingness + scale(month_index) + I(scale(month_index)^2) + (1 | party) +
                         offset(log(total_nr_sentences)),
                       family = nbinom1,
                       data = monthly_blame)

summary(model_bloc_wing)
diagnose(model_bloc_wing)
```


```{r}
model_bloc_wing_int <- glmmTMB(total_blame_corrected ~ in_gov + bloc * wingness + scale(month_index) + I(scale(month_index)^2) + (1 | party) +
                             offset(log(total_nr_sentences)),
                           family = nbinom1,
                           data = monthly_blame)

summary(model_bloc_wing_int)
```

```{r}
model_bloc_bloc_wing_int <- glmmTMB(total_blame_corrected ~ in_gov*bloc + bloc * wingness + scale(month_index) + I(scale(month_index)^2) + (1 | party) +
                             offset(log(total_nr_sentences)),
                           family = nbinom1,
                           data = monthly_blame)

summary(model_bloc_bloc_wing_int)
```
```{r}
exp(0.553758)
```



```{r}
anova(model_intercept,model_in_gov,model_bloc,model_bloc_wing,model_bloc_wing_int, model_bloc_bloc_wing_int)
```

```{r}
anova(model_in_gov, model_bloc_wing, model_bloc_wing_int, model_bloc_bloc_wing_int)
```

```{r}
anova(model_in_gov, model_bloc_wing_int, model_bloc_bloc_wing_int)
```

#check assumptions of best model

```{r}
model_check(model_bloc_bloc_wing_int)
```


```{r}
summary(model_bloc_bloc_wing_int)
```




#plot it

```{r}
model_bloc_bloc_wing_int <- glmmTMB(total_blame_corrected ~ in_gov*bloc + bloc * wingness + month_index_scaled + month_index_scaled_sqrt + (1 | party) +
                             offset(log(total_nr_sentences)),
                           family = nbinom1,
                           data = monthly_blame)

summary(model_bloc_bloc_wing_int)
```



```{r}
sjPlot::plot_model(model_bloc_bloc_wing_int, 
                   type = "pred", 
                   terms = c("in_gov", "bloc"), 
                   condition = c(total_nr_sentences = 100),
                   axis.title = c("In government", "Predicted Blame Counts pr 100 Uttered Sentences"),
                   title = "Blame Predicted by Government status and ideological bloc",
                   #ci.lvl = NA,
                   #axis.lim = c(3,12)
                   ) +
  theme_minimal()


sjPlot::plot_model(model_bloc_bloc_wing_int, 
                   type = "pred", 
                   terms = c("wingness"), 
                   condition = c(total_nr_sentences = 100),
                   axis.title = c("Wingness", "Predicted Blame Counts pr 100 Uttered Sentences"),
                   title = "Blame Predicted by Government status and ideological bloc",
                   #ci.lvl = NA,
                   #axis.lim = c(3,12)
                   ) +
  theme_minimal()

sjPlot::plot_model(model_bloc_bloc_wing_int, 
                   type = "pred", 
                   terms = c("bloc","wingness"), 
                   condition = c(total_nr_sentences = 100),
                   axis.title = c("Bloc", "Predicted Blame Counts pr 100 Uttered Sentences"),
                   title = "Blame Predicted by Government status and ideological bloc",
                   #ci.lvl = NA,
                   #axis.lim = c(3,12)
                   ) +
  theme_minimal()
```



# DO check to see if quadratic control variable for time alters focal predicters


```{r}
# Compare linear vs quadratic time control in Family 2
model_block_int_linear <- glmmTMB(total_blame_corrected ~ in_gov*bloc + bloc * wingness + month_index + (1 | party) +
                             offset(log(total_nr_sentences)),
                           family = nbinom1,
                           data = monthly_blame)

model_block_int_quadratic <- glmmTMB(total_blame_corrected ~ in_gov*bloc + bloc * wingness + scale(month_index) + I(scale(month_index)^2) + (1 | party) +
                             offset(log(total_nr_sentences)),
                           family = nbinom1,
                           data = monthly_blame)

AIC(model_block_int_linear, model_block_int_quadratic)

anova(model_block_int_linear, model_block_int_quadratic)
```


```{r}
# Also check whether political coefficients change materially
summary(model_block_int_linear)

```


```{r}
summary(model_block_int_quadratic)
```



# For newer times


#Make data

```{r}
monthly_blame_18 <- monthly_blame %>% 
  filter(year >=19) %>% 
  filter(year <=26) %>% 
  mutate(month_index = month_index-255)
```




#Do analysis
monthly_blame_18
`
```{r}
model_intercept_18 <- glmmTMB(total_blame_corrected ~ 1 + month_index + (1|party)+ offset(log(total_nr_sentences)), 
                           family = nbinom1, 
                           data = monthly_blame_18)

summary(model_intercept_18)
```

```{r}
model_in_gov_18 <- glmmTMB(total_blame_corrected ~ in_gov + month_index + (1 | party) +
                          offset(log(total_nr_sentences)),
                        family = nbinom1,
                        data = monthly_blame_18)

summary(model_in_gov_18)
```



```{r}
model_bloc_18 <- glmmTMB(total_blame_corrected ~ in_gov + bloc + month_index + (1 | party) +
                         offset(log(total_nr_sentences)),
                       family = nbinom1,
                       data = monthly_blame_18)

summary(model_bloc_18)
```

```{r}
model_bloc_wing_18 <- glmmTMB(total_blame_corrected ~ in_gov + bloc + wingness + month_index + (1 | party) +
                         offset(log(total_nr_sentences)),
                       family = nbinom1,
                       data = monthly_blame_18)

summary(model_bloc_wing_18)
```

```{r}
model_bloc_wing_int_18 <- glmmTMB(total_blame_corrected ~ in_gov + bloc * wingness + month_index + (1 | party) +
                             offset(log(total_nr_sentences)),
                           family = nbinom1,
                           data = monthly_blame_18)

summary(model_bloc_wing_int_18)
```


```{r}
model_in_gov_bloc_wing_int_18 <- glmmTMB(total_blame_corrected ~ in_gov *bloc + bloc * wingness + month_index + (1 | party) +
                             offset(log(total_nr_sentences)),
                           family = nbinom1,
                           data = monthly_blame_18)

summary(model_in_gov_bloc_wing_int_18)
```




```{r}
anova(model_intercept_18, model_in_gov_18, model_bloc_18, model_bloc_wing_18, model_bloc_wing_int_18, model_in_gov_bloc_wing_int_18)
```

```{r}
anova(model_in_gov_18, model_bloc_wing_int_18, model_in_gov_bloc_wing_int_18)
```



```{r}
anova(model_in_gov_18, model_bloc_wing_int_18)
```

```{r}
#best model
summary(model_bloc_wing_int_18)
```
```{r}
exp(0.8171468)
```

#check assumption of best model

```{r}
model_check(model_bloc_wing_int_18)
```
#Plot best model

```{r}
sjPlot::plot_model(model_bloc_wing_int_18, 
                   type = "pred", 
                   terms = c("in_gov", "bloc"), 
                   condition = c(total_nr_sentences = 100),
                   axis.title = c("In Government", "Predicted Blame Counts pr 100 Uttered Sentences"),
                   title = "Predicted Blame by In Government and ideological Bloc in period 2019-2022",
                   #ci.lvl = NA,
                   #axis.lim = c(3,12)
                   ) +
  theme_minimal()
sjPlot::plot_model(model_bloc_wing_int_18, 
                   type = "pred", 
                   terms = c("in_gov"), 
                   condition = c(total_nr_sentences = 100),
                   axis.title = c("In Government", "Predicted Blame Counts pr 100 Uttered Sentences"),
                   title = "Predicted Blame by In Government and ideological Bloc in period 2019-2022",
                   #ci.lvl = NA,
                   #axis.lim = c(3,12)
                   ) +
  theme_minimal()

sjPlot::plot_model(model_bloc_wing_int_18, 
                   type = "pred", 
                   terms = c("bloc"), 
                   condition = c(total_nr_sentences = 100),
                   axis.title = c("Bloc", "Predicted Blame Counts pr 100 Uttered Sentences"),
                   title = "Predicted Blame by In Government and ideological Bloc in period 2019-2022",
                   #ci.lvl = NA,
                   #axis.lim = c(3,12)
                   ) +
  theme_minimal()

sjPlot::plot_model(model_bloc_wing_int_18, 
                   type = "pred", 
                   terms = c("wingness"), 
                   condition = c(total_nr_sentences = 100),
                   axis.title = c("Wingness", "Predicted Blame Counts pr 100 Uttered Sentences"),
                   title = "Predicted Blame by Wing status in period 2019-2022",
                   #ci.lvl = NA,
                   #axis.lim = c(3,12)
                   ) +
  theme_minimal()

sjPlot::plot_model(model_bloc_wing_int_18, 
                   type = "pred", 
                   terms = c("bloc","wingness"), 
                   condition = c(total_nr_sentences = 100),
                   axis.title = c("bloc", "Predicted Blame Counts pr 100 Uttered Sentences"),
                   title = "Predicted Blame by Wing status in period 2019-2022",
                   #ci.lvl = NA,
                   #axis.lim = c(3,12)
                   ) +
  theme_minimal()
```



#time as predictor of blame in recent years

```{r}
model_blame_time_18_intercept <- glmmTMB(total_blame_corrected ~ 1 + in_gov  + (1 | party) + offset(log(total_nr_sentences)), 
                           family = nbinom1, 
                           data = monthly_blame_18)

summary(model_blame_time_18_intercept)
```

```{r}
model_blame_time_18 <- glmmTMB(total_blame_corrected ~  in_gov + scale(month_index)  + (1 | party) + offset(log(total_nr_sentences)), 
                           family = nbinom1, 
                           data = monthly_blame_18)

summary(model_blame_time_18)
```

```{r}
model_blame_time_18_2 <- glmmTMB(total_blame_corrected ~ in_gov + scale(month_index) + I(scale(month_index)^2)  + (1 | party) + offset(log(total_nr_sentences)), 
                           family = nbinom1, 
                           data = monthly_blame_18)

summary(model_blame_time_18_2)
```
```{r}
anova(model_blame_time_18_intercept,model_blame_time_18, model_blame_time_18_2)
```

```{r}
model_check(model_blame_time_18)
```

# OBS due to convergence issues/singular fit, nbinom 2 was needed here

```{r}
#best model
summary(model_blame_time_18)
```
```{r}
exp(0.092078)
```

#Plot this model

```{r}
sjPlot::plot_model(model_blame_time_18, 
                   type = "pred", 
                   terms = c("month_index"), 
                   condition = c(total_nr_sentences = 100),
                   axis.title = c("Month since Jan 2019", "Predicted Blame Counts pr 100 Uttered Sentences"),
                   title = "Evolution of the Use of Blame in period 2019-2022",
                   #ci.lvl = NA,
                    #axis.lim = c(3,14)
                   ) +
  theme_minimal()
```

```{r}
sum(monthly_blame$total_blame_corrected)/sum(monthly_blame$total_nr_sentences)
```


```{r}
table(monthly_blame$bloc, monthly_blame$wingness)
```



